---
title: "The title goes here"
author: "Lemon Reimer"
date: "3/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Ignore this part
```

# This is a title

### This is a subtitle

This is a very **bold** and *slanted* sentence

This looks a lot like `code`.

```{r your first code block}
# this is your first code block, where code can be written
# anything after the '#' is a comment and will not run
# try treating this section as a calculator. type 2 + 2, then click the green arrow
```

# Libraries and importing data

```{r}
library(dplyr)
#library(DT)

# load the 'datasets' package
dplyr::filter()
```

```{r}
test_data <- read.csv("https://vincentarelbundock.github.io/Rdatasets/csv/Ecdat/Caschool.csv")

# import the data from https://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-red.csv and call it wine_red

# import the data from https://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv and call it wine_white
```

# Subsetting data by columns

```{r}
col_keep <- c("distcod", "enrltot", "teachers", "calwpct", "mealpct", "computer", "testscr", "compstu", "expnstu", "str", "avginc", "elpct", "readscr", "mathscr")
col_rmv <- c("X", "county", "district", "grspan")
```

### Subset using wanted column names

```{r}
sub1 <- test_data[col_keep]
```

```{r}
sub2 <- subset(test_data, select = col_keep)
```

```{r}
sub3 <- dplyr::select(test_data, col_keep)
```

### Subset using unwanted column names

```{r}
sub4 <- test_data[!names(test_data) %in% col_rmv]
!names(test_data) %in% col_rmv
```

```{r}
sub5 <- subset(test_data, select = -c(X, county, district, grspan))
```

```{r}
sub6 <- select(test_data, -col_rmv)
```

# Subsetting data by boolean

```{r}
sub7 <- test_data[test_data$county == "Butte", col_keep]
```

```{r}
sub8 <- subset(test_data, county == "Butte", select = col_keep)
```

```{r}
sub9 <- select(filter(test_data, county == "Butte"), col_keep)
```

# Deleting/removing data

```{r}
# Experiment, try to figure out how to remove the "X" column
# If you get stuck, go back up and re-run where we import the data to refresh the dataset
# If that was too easy, try to remove rows based on a boolean
```

# Replacing data

```{r}
# Experiment again! Think to the last section and see what we should change about how we've been writing our code


```

# Calculating a new column

### Dummy variable

```{r}
test_data$dummy <- ifelse(test_data$grspan == "KK-08", 1, 0)
table(test_data$dummy, test_data$grspan)
```

```{r}
test_data$cats <- ifelse((test_data$county == "Butte"), 1, ifelse((test_data$county == "Fresno"), 2, 0))
```

### Composite scoring

```{r}
# The column 'avginc' is an average of readscr and mathscr
# Try to re-create this variable in a new column called 'avgs'

test_data$avgs <- (test_data$mathscr + test_data$readscr)/2

test_data$avgs <- test_data$mathscr + test_data$readscr

test_data$avgs <- test_data$avgs/2
```

```{r}
# Look up 'mutate' in the help tab and read up to "Useful functions available"
```

# Combining datasets

### Adding rows

```{r}
# POP QUIZ!!
# Add a new column to the wine_red and wine_white datasets called "color"
# In wine_white, "color" = 1 and in wine_red, "color" = 0
```

```{r}
wines <- rbind(wine_red, wine_white)
```

```{r}
wines <- merge(wine_red, wine_white, all = T)
```

### Adding columns

```{r}
# You can ignore this if you want, I'm just building two datasets with random ACT/SAT scores

df_ACT <- data.frame("id" = c(1:20), 
                     "ACT_math" = c(floor(runif(20, min = 1, max = 36))), 
                     "ACT_english" = c(floor(runif(20, min = 1, max = 36))), 
                     "ACT_science" = c(floor(runif(20, min = 1, max = 36))), 
                     "ACT_reading" = c(floor(runif(20, min = 1, max = 36))))
df_ACT$ACT_avg <- (df_ACT$ACT_math + df_ACT$ACT_english + df_ACT$ACT_science + df_ACT$ACT_reading)/4

df_SAT <- data.frame("id" = c(1:20), 
                     "SAT_math" = c(floor(runif(20, min = 200, max = 800))), 
                     "SAT_reading" = c(floor(runif(20, min = 200, max = 800))))
df_SAT$SAT_total <- df_SAT$SAT_math + df_SAT$SAT_reading

```

```{r}
df_SCORE <- cbind(df_ACT, df_SAT)
```

```{r}
df_SCORE <- merge(df_ACT, df_SAT, by = "id")
```

```{r}
df_SCORE <- full_join(df_ACT, df_SAT, by = "id")
```

# Dplyr pipelines

```{r}
# this is the same thing as select(test_data, col_keep) but in a pipeline

sub3 <- test_data %>% select(col_keep)
```

```{r}
# Try bulding a filter pipeline
```

```{r}
# Try building a mutate pipeline
```

# Attaching a dplyr pipeline to DT

```{r}
# We'll do this together!
```

